Species richness and network structure jointly drive total biomass and its temporal stability in fish communities

### 2020-10-22 manuscript/bef_stability/manuscript.Rmd
Add affiliations

- Author affiliation:
  - Alain Danet:
  - Maud Mouchet:
  - Willem Bonnaffe: Ecological and Evolutionary Dynamics Lab. Zoology Research
    and Administration Building, 11a Mansfield Rd, Oxford OX1 3SZ, United
    Kingdom.
  - Elisa Thebault: 
  - Colin Fontaine:

@@

- Number of words:
  - Abstract: 184
  - Main text: 4929

@@

- Coresponding author:
  - Alain Danet
  - UMR 7204 CESCO, BP 61; 45, rue Buffon, 75005 Paris, France
  - alain.danet@mnhn.fr 
  - +33140798134

@@

Both the relationship between diversity and ecosystem functioning, and between food
web complexity and stability, have received considerable attention during the
last decades in ecology. The former relationship remains rarely investigated
in the context of food webs, especially in empirical studies explicitly
considering food web structures, the latter has been mainly the focus of
theoretical studies and empirical tests are critically missing. Here we link
species richness, food web structure and annual biomass and its temporal stability
using a dataset of standardized fish sampling in French streams running from
1995 to 2018. We inferred trophic interactions thanks to recorded body size and
diet literature. We reveal that species richness and average trophic level had a
positive relationship with annual biomass of fish communities while they had
respectively a negative and positive effect on its temporal stability. We
also found that the effects of stream size and temperature on annual biomass and
stability are partly mediated by their effects on diversity and food web
structure. Overall, our study opens fruitful perspectives on the link between
species richness, network structure, ecosystem properties and their stability.

@@

Species richness promotes higher biomass production in a variety of ecosystems
and taxons. Indeed species richness generally increases primary productivity and
total biomass in experimental plant communities [@tilman_productivity_1996;
@grace_integrative_2016; @hector_plant_1999; @tilman_biodiversity_2014;
@grace_integrative_2016], but also in briophyte communities
[@mulder_physical_2001], stream feeders, [@cardinale_species_2002] and aquatic
bacterial consumers [@pennekamp_biodiversity_2018]. Despite the diversity of
ecosystems and taxons studied, the relationship between diversity and ecosystem
functioning remains traditionally focused on single trophic levels, ignoring the
effects of food web structure. Theoretical models have recently moved from
single trophic [@tilman_plant_1997;
@loreau_biodiversity_2001] to multitrophic communities [@thebault_food-web_2003;
@poisot_dissimilarity_2012; @schneider_animal_2016; @wang_biodiversity_2018]. 
These models show that both resource use complementarity and selection effects,
the two main mechanisms previously highlighted in single-trophic studies, are
also important to understand diversity effects in food webs
[@thebault_food-web_2003; @poisot_dissimilarity_2012; @schneider_animal_2016].
However, they also reveal new mechanisms such as positive effects of 
the number of trophic levels, i.e. associated with higher trophic levels, on ecosystem
functioning [@wang_biodiversity_2018]. Theoretical studies also showed that food
web properties, such as interaction distribution and strength, affect the
distribution of biomass among adjacent trophic levels
[@thebault_relationship_2006; @schneider_animal_2016; @barbier_pyramids_2019].
Despite these theoretical evidences that structure affects biomass, we are
lacking empirical studies that assess jointly how food web structure and species
diversity affect total biomass and its production.

@@

Further than biomass production, many studies showed that species richness
increases temporal stability of communities, i.e. variation through time, often
quantified by the temporal CV of biomass.  Such positive relationship between
species richness and temporal stability of biomass has been reported in
experimental grasslands and aquatic bacterial systems
[@tilman_biodiversity_2006; @pennekamp_biodiversity_2018] as well as in natural
settings, such as stream fish communities [@franssen_annual_2011], bird and
butterfly communities [@olivier_urbanization_2020]. However, this relationship
has been mostly explored in single trophic level systems or without accounting
for food web structure, even more so than the relationship between diversity and
biomass production.

@@

In addition to diversity,
food web structure is expected to strongly affect the stability of communities
and ecosystems, as outlined by the large body of theoretical studies on this
topic [@may_will_1972; @dunne_network_2006; @neutel_reconciling_2007;
@duffy_why_2009; @stouffer_compartmentalization_2011]. In particular, predation
has cascading effects on both species variability and synchrony
[@teng_dynamics_2004; @shanafelt_stability_2018] and it also affects
synchrony between species.

@@

Overall, the evidence suggests that both species richness and network structure
should modulate total biomass and stability of communities. The effects of
species richness and network structure, such as connectance and average trophic
level, are potentially conflicting or synergistic in such a way that they need
to be addressed jointly. Despite the long-standing interest for the links
between food web complexity and stability in ecology, we still have very few
evidence of the relationship between network structure and the temporal
stability of multitrophic communities in natural settings.
Although the relationships between diversity, total biomass and its stability
found in experimental and in natural settings can be similar
[e.g., @wu_relationship_2015; @li_relationship_2018], negative or non
significant effects of species richness on total biomass have been found in some
natural communities such as forests [@szwagrzyk_above-ground_2007;
@jacob_productivity_2010; @seidel_relationship_2013]. Some authors suggest that
the lack of resource complementarity or environmental harshness might be
responsible for negligible, or negative effects of species richness
[@loreau_biodiversity_2001; @he_global_2013; @wu_relationship_2015].

@@

Our aim is to document the relationships among species richness, food web structure, 
total biomass and its temporal stability in natural communities. To do so, we
used the ONEMA dataset consisting in standardized sampling of fish communities
in French streams [@edeline_ecological_2013]. We selected 99 sites that were sampled for more than 10
years spanning on the 1995-2018 period. We inferred 1496 interaction networks,
based on body size and diet. We question whether species richness, connectance
and average trophic level increase, decrease or are not related to
total biomass and its temporal stability. We further question whether the
species richness, connectance, and average trophic level have antagonistic or
mutual effect on total biomass and its temporal stability. Finally, we
question if ecosystem size and average temperature increase total biomass and
its temporal stability.

@@

```{r}
matmet_cap <- paste0(
  #"(A) Conceptual model of the hypothesized relationships between environment,
  #species richness, network metrics and total biomass / biomass stability. ",
  #"Solid and dashed arrows represent respectively direct and indirect effects.
  #",
  #"The blue, green and blue/green dashed arrows being respectively indirect
  #effects going through species richness, network metrics and through both
  #species richness and network metrics.",
  "Yearly temporal series of networks of two stations having respectively a low (A) and high
  (B) temporal stability of biomass. A and B stations have respectively the
  closest stability to the first and third quartile of the distribution of the
  temporal stability of biomass in the dataset. Point size are proportional to
  the biomass of the trophic species.  (C) Observation sites (black dots) were
  located across France and monitored on the period 1995-2018. ", "Grey shades
  highlighted elevalation variation. ", "Inner borders draw the limits of the
  hydrographic basins. ", "(D) The metaweb. Each trophic species (N = 412) is
  represented by a node and each color represent a species (S = 45). Resource
  nodes (7) were drawn in grey. ", "The species names corresponding to the three
  digit codes are found in Table SXX."
)
```

@@

# Select fishing operation 
myload(op_analysis, op_analysis_wo_holes, dir = mypath("data"))
myload(biomass_ts_sax, dir = mypath("data"))
op_analysis_bbb <- filter(op_analysis, station %in%
  biomass_ts_sax[biomass_ts_sax$sax == "bbb",]$station)
op_analysis_wo_holes_bbb <- filter(op_analysis_wo_holes, station %in%
  biomass_ts_sax[biomass_ts_sax$sax == "bbb",]$station)
range_surface <- map_dbl(c(min, max), ~.x(op_analysis_bbb$surface, na.rm =TRUE))
# quantile(op_analysis_bbb$surface, na.rm =TRUE)
```

@@

Structural Equation Modeling enables us to assess multivariate relationships
between variables. It is a promising way to go beyond bivariate relationships
and to investigate complex relationships, direct and
indirect, between variables [@grace_structural_2008; @grace_causal_2014;
@lefcheck_piecewisesem_2016; @grace_integrative_2016]. The structural equation
model explaining the annual total biomass allowed the environment
to affect directly species richness and network structure, as well as annual total
biomass (Figure \@ref(fig:sem-plot) B). We assumed that both species richness
and network structure directly affect total biomass (Figure \@ref(fig:sem-plot)
B) and that species richness affected food web structure
[@winemiller_must_1989; @dunne_network_2006; @thebault_stability_2010].

@@

library(DiagrammeR)
meta_stab_sem_rich <- export_diagram(
  fit = stab_sem_rich,
  file_path = mypath("manuscript", "bef_stability", "meta_stab_sem"),
  format = "png",
  width = 700*2,
  height = 700*2 * asp,
  arrows = "ortho",
  val_thd = 0.05,
  env_x_pos_range = def_env_x_pos,
  env_y_pos = 0,
  env_y_sep = 1.5,
  order_env_node = c("log_RC1", "log_RC2"),
  metamodel = TRUE,
  inverse_y_node_pose = TRUE
)
p_stab_sem_rich <- export_diagram(
  fit = stab_sem_rich,
  file_path = mypath("manuscript", "bef_stability", "stab_sem"),
  format = "png",
  width = 700*2,
  height = 700*2 * asp,
  arrows = "ortho",
  val_thd = 0.05,
  env_x_pos_range = def_env_x_pos,
  env_y_pos = 0,
  env_y_sep = 1.5,
  order_env_node = c("log_RC1", "log_RC2"),
  inverse_y_node_pose = TRUE
)

meta_bm_sem_rich <- export_diagram(
  fit = bm_sem_rich,
  file_path = mypath("manuscript", "bef_stability", "meta_bm_sem"),
  format = "png",
  width = 700*2,
  height = 700*2 * asp,
  arrows = "ortho",
  val_thd = 0.05,
  env_x_pos_range = def_env_x_pos,
  env_y_pos = 0,
  env_y_sep = 1.5,
  order_env_node = c("log_RC1", "log_RC2"),
  metamodel = TRUE,
  inverse_y_node_pose = TRUE
)
p_bm_sem_rich <- export_diagram(
  fit = bm_sem_rich,
  file_path = mypath("manuscript", "bef_stability", "bm_sem"),
  format = "png",
  width = 700*2,
  height = 700*2 * asp,
  arrows = "ortho",
  val_thd = 0.05,
  env_x_pos_range = def_env_x_pos,
  env_y_pos = 0,
  env_y_sep = 1.5,
  order_env_node = c("log_RC1", "log_RC2"),
  inverse_y_node_pose = TRUE
)
```

@@

```{r}
sem_caption <- paste0(
  "Metamodel of the Structural Equation Models for (A) biomass stability and
  (B) total biomass. ",
  "Results of the Structural Equation Models for (C) the
  determinants of biomass stability and (D) total biomass. Red and green
  represent respectively negative and positive path coefficients. The path
  coefficients are standardized coeffients. For convenience, only significant
  paths (i.e, *P < 0.05*) are shown. Rsq: Ajusted R square."

)
```

```{r sem-plot, fig.height = 7, fig.cap = sem_caption}
#https://tex.stackexchange.com/questions/20496/drawing-different-tikz-shapes-parameterized-by-data-from-a-file?rq=1
#https://tex.stackexchange.com/questions/83888/how-to-plot-data-from-a-csv-file-using-tikz-and-csvsimple

@@

```{r indir-stab}
caption_sem <- paste0(
  "Direct, indirect and total effect of environment, species richness and
  network structure, i.e. connectance and mean trophic level on total biomass
  and its temporal stability. Indirect effects are computed by multiplying the
  path coefficients from the variable predictor to the response variable, here
  biomass stability. Indirect effects are considered as one step from the target
  variable. For exemple, the indirect effect of species richness on
  biomass stability via its effect on CVsp is $-0.61$. To compute indirect
  effect, only the significant direct effects were taken into account. ",
  "As an example, the indirect effect via species richness takes into account
 the direct effect of species richness as well as the effect of species richness
 through connectance and average trophic level. ",
 "NA values indicate that the corresponding paths do not exist in the
 specified Structural Equation Model (SEM)."
)

@@

library(kableExtra)
mytable <- kable(full_table,
  format = "latex",
  booktabs = T,
  label = paste0("indir-sem"),
  caption = caption_sem
  ) %>%
#kable_paper() %>%
kable_styling(latex_options =c("scale_down")) %>%
column_spec(1:2, width = "3cm") %>%
collapse_rows(columns = 1, valign = "top")
write(mytable, file = mypath("manuscript", "bef_stability", "table.tex"))
```

@@

Surprisingly we found a negative relationship between temporal stability and
species richness. Overall, it contrasts with experimental studies in grasslands
[@tilman_biodiversity_2006]; mesocosms (Pennekamp et al. 2018), in 
butterflies natural communities [@olivier_urbanization_2020] or even in lotic
fish communities [@franssen_annual_2011] where positive relationships were found
between temporal stability and species richness. However, other studies found an
opposite relationship in natural grassland [@yang_effects_2011] and a marginal
one in natural forests [@declerck_species_2006].

A first step to identify the mechanisms behind those contrasting patterns is to
look at the effects of species richness on synchrony and $\overline{CV_{sp}}$.
As most studies on the topic, in experimental grasslands
[@isbell_biodiversity_2009; @roscher_identifying_2011], in natural communities
of birds, grasslands, bats and butterflies [@bluthgen_land_2016;
@olivier_urbanization_2020], we found that species richness decreases synchrony,
thereby increasing stability [but see @declerck_species_2006]. Our study thus
extends the evidence that species richness decrease synchrony to the case of empirical
trophic communities. Species richness can reduce synchrony when a given
increase in species richness leads to a decrease of population sizes but also in
the case where species have different responses to environmental variation
[@loreau_species_2008].

The positive effect of species richness on
$\overline{CV_{sp}}$, thereby decreasing stability, is more intriguing. While
positive effect of species richness on $\overline{CV_{sp}}$ was reported in
experimental grasslands and natural bird communities [@tilman_biodiversity_2006;
@olivier_urbanization_2020], negative and neutral relationships have been found
in butterfly and bat communities. Our results contrast with a previous
meta-analysis which concludes that species richness had a positive effect on
population stability in multi-trophic aquatic microcosms
[@jiang_different_2009]. In foodwebs, the relationship between temporal
stability of populations and species richness depends on interaction strength
and their distribution [@mccann_reevaluating_1997; @thebault_trophic_2005;
@brose_allometric_2006], high interaction strength inducing a negative
relationship. Therefore, it is possible that our results are due to a high
degree of omnivory in food webs, which combined with strong trophic interactions
lead to more fluctuations in population biomass.

@@

In contrast, connectance had no effect on total biomass. One might have expected
that an increase in connectance may translate in higher total nutrient uptake
and so results in higher total biomass in the fish compartment. A reason for
such an absence of an effect might be explained by the fact that our data lack
information about the lower compartments of the community, i.e, the primary
producers, zooplankton and zoobenthos.

Mean trophic level had a positive effect on biomass stability by its negative
effect on $\overline{CV_{sp}}$, thereby increasing average species stability. It
might be because species at higher trophic level are more stable than the species at
lower trophic levels as predicted by food chain models in the presence of
strong top-down control [@barbier_pyramids_2019; @shanafelt_stability_2018] but it
is not sure that it holds when considering omnivorous species
[@wang_biodiversity_2016]. An alternative explanation is that higher mean
trophic level leads to stronger top-down control and resource complementarity
and thus reduce biomass fluctuations in lower trophic levels
[@rooney_structural_2006].

@@

In contrast with previous studies, we found that the average population
variability ($\overline{CV_{sp}}$) had a stronger negative effect on biomass
stability than synchrony [@tilman_biodiversity_2006; @bluthgen_land_2016;
@olivier_urbanization_2020]. It explains why we observe a negative effect of
species richness on temporal stability of biomass. The lower effect of synchrony
might be explained by the presence of generalist predators in stream
communities, which can raise synchrony between prey species which share the
same predator [@raimondo_interspecific_2004] and thus decreases the
occurrence of compensatory population dynamics. Alternatively, it is possible that, in
aquatic foodwebs, the occurrence of large oscillatory dynamics
[@mccann_reevaluating_1997] leads to a higher contribution of population
variability rather than synchrony to community stability. How both
strengths and the relative contribution of synchrony and population
variaribility to overall stability vary between foodwebs and single trophic communities is
still an open question.

@@

Consistent with research showing that food webs in wider streams, are also
larger, and hence support more species and biomass [@allan_stream_2007], we
found that the PCA axis related to stream size had a positive effect on total
biomass through a positive effect on mean trophic level. Greater ecosystem size
might provide a greater diversity of resources through habitat heterogeneity,
and so higher species richness and total biomass by surface unit, through
resource complementarity. Greater diversity of resources can also maintain
more biomass at high trophic level [@post_ecosystem_2000; @doi_resource_2009].
Furthermore, the positive effect of the PCA axis related to temperature on total
biomass was of the magnitude of effects of network structure and species
richness, in line with the mounting body of work reporting important effects of
temperature on the structure and biomass of aquatic system [e.g.
@dossena_warming_2012; @edeline_ecological_2013; @emmrich_geographical_2014;
@hattab_forecasting_2016]. Contrasting with a previous study in experimental
plant communities [@grace_integrative_2016] and in fish communities
[@yvondurocher_warming_2011; @dossena_warming_2012], we found that environmental
drivers overall affect total biomass indirectly rather than directly.

>>>>

- Author affiliation:
  - Alain Danet: Centre d’Ecologie et des Sciences de la Conservation, UMR 7204
    MNHN-CNRS-Sorbonne Université, Muséum national d’Histoire naturelle de
    Paris, 43 rue Buffon, 75005 Paris, France
  - Maud Mouchet: Centre d’Ecologie et des Sciences de la Conservation, UMR 7204
    MNHN-CNRS-Sorbonne Université, Muséum national d’Histoire naturelle de
    Paris, 43 rue Buffon, 75005 Paris, France
  - Willem Bonnaffe: Ecological and Evolutionary Dynamics Lab. Zoology Research
    and Administration Building, 11a Mansfield Rd, Oxford OX1 3SZ, United
    Kingdom.
  - Elisa Thebault: Sorbonne Université, CNRS, IRD, INRAE, Université Paris Est
    Créteil, Institute of Ecology and Environmental Sciences of Paris
    (iEES-Paris), Paris, France  
  - Colin Fontaine: Centre d’Ecologie et des Sciences de la Conservation, UMR 7204
    MNHN-CNRS-Sorbonne Université, Muséum national d’Histoire naturelle de
    Paris, 43 rue Buffon, 75005 Paris, France

@@

- Number of words:
  - Abstract: 172
  - Main text: 4929

@@

- Coresponding author:
  - Alain Danet
  - UMR 7204 CESCO, CP 135; 43, rue Buffon, 75005 Paris, France
  - alain.danet@mnhn.fr 
  - +33140798134

@@

Both biodiversity-ecosystem functioning (BEF) and foodweb complexity-stability
relationships have received considerable attention during the last decades.
Still, the BEF relationship remains rarely investigated in the context of food
webs, especially in empirical studies explicitly considering foodweb structures.
Foodweb stability-complexity relationship has been mainly the focus of
theoretical studies with very empirical testing. To fill this gap, we link
species richness, foodweb structure and annual biomass and its temporal
stability using a standardized monitoring of fish communities in French streams
running from 1995 to 2018. We inferred trophic interactions thanks to recorded
body size and diet literature. We reveal that species richness and average
trophic level had a positive relationship with annual biomass of fish
communities while they had respectively a negative and positive effect on its
temporal stability. We also found that the effects of stream size and
temperature on annual biomass and stability are mediated by their effects on
diversity and foodweb structure. Overall, our study opens fruitful perspectives
on the link between species richness, network structure, biomass and its
stability.

@@

Species richness promotes higher biomass production in a variety of ecosystems
and taxons. Indeed species richness generally increases primary productivity and
total biomass in experimental plant communities [@tilman_productivity_1996;
@grace_integrative_2016; @hector_plant_1999; @tilman_biodiversity_2014;
@grace_integrative_2016], but also in briophyte communities
[@mulder_physical_2001], stream feeders, [@cardinale_species_2002] and aquatic
bacterial consumers [@pennekamp_biodiversity_2018]. Despite the diversity of
ecosystems and taxons studied, the relationship between diversity and ecosystem
functioning remains traditionally focused on single trophic levels, ignoring the
effects of foodweb structure. Theoretical models have recently moved from
single trophic [@tilman_plant_1997;
@loreau_biodiversity_2001] to multitrophic communities [@thebault_food-web_2003;
@poisot_dissimilarity_2012; @schneider_animal_2016; @wang_biodiversity_2018]. 
These models show that both resource use complementarity and selection effects,
the two main mechanisms previously highlighted in single-trophic studies, are
also important to understand diversity effects in foodwebs
[@thebault_food-web_2003; @poisot_dissimilarity_2012; @schneider_animal_2016].
However, they also reveal new mechanisms such as positive effects of 
the number of trophic levels, i.e. associated with higher trophic levels, on ecosystem
functioning [@wang_biodiversity_2018]. Theoretical studies also showed that food
web properties, such as interaction distribution and strength, affect the
distribution of biomass among adjacent trophic levels
[@thebault_relationship_2006; @schneider_animal_2016; @barbier_pyramids_2019].
Despite these theoretical evidences that structure affects biomass, we are
lacking empirical studies that assess jointly how foodweb structure and species
diversity affect total biomass and its production.

@@

Further than biomass production, many studies showed that species richness
increases temporal stability of communities, i.e. variation through time, often
quantified by the temporal CV of biomass.  Such positive relationship between
species richness and temporal stability of biomass has been reported in
experimental grasslands and aquatic bacterial systems
[@tilman_biodiversity_2006; @pennekamp_biodiversity_2018] as well as in natural
settings, such as stream fish communities [@franssen_annual_2011], bird and
butterfly communities [@olivier_urbanization_2020]. However, this relationship
has been mostly explored in single trophic level systems or without accounting
for foodweb structure, even more so than the relationship between diversity and
biomass production.

@@

In addition to diversity,
foodweb structure is expected to strongly affect the stability of communities
and ecosystems, as outlined by the large body of theoretical studies on this
topic [@may_will_1972; @dunne_network_2006; @neutel_reconciling_2007;
@duffy_why_2009; @stouffer_compartmentalization_2011]. In particular, predation
has cascading effects on both species variability and synchrony
[@teng_dynamics_2004; @shanafelt_stability_2018] and it also affects
synchrony between species.

@@

Overall, the evidence suggests that both species richness and network structure
should modulate total biomass and stability of communities. The effects of
species richness and network structure, such as connectance and average trophic
level, are potentially conflicting or synergistic in such a way that they need
to be addressed jointly. Despite the long-standing interest for the links
between foodweb complexity and stability in ecology, we still have very few
evidence of the relationship between network structure and the temporal
stability of multitrophic communities in natural settings.
Although the relationships between diversity, total biomass and its stability
found in experimental and in natural settings can be similar
[e.g., @wu_relationship_2015; @li_relationship_2018], negative or non
significant effects of species richness on total biomass have been found in some
natural communities such as forests [@szwagrzyk_above-ground_2007;
@jacob_productivity_2010; @seidel_relationship_2013]. Some authors suggest that
the lack of resource complementarity or environmental harshness might be
responsible for negligible, or negative effects of species richness
[@loreau_biodiversity_2001; @he_global_2013; @wu_relationship_2015].

@@

Our aim is to document the relationships among species richness, foodweb structure, 
total biomass and its temporal stability in natural communities. To do so, we
used the ONEMA dataset consisting in standardized sampling of fish communities
in French streams [@edeline_ecological_2013]. We selected 99 sites that were sampled for more than 10
years spanning on the 1995-2018 period. We inferred 1496 interaction networks,
based on body size and diet. We question whether species richness, connectance
and average trophic level increase, decrease or are not related to
total biomass and its temporal stability. We further question whether the
species richness, connectance, and average trophic level have antagonistic or
mutual effect on total biomass and its temporal stability. Finally, we
question if ecosystem size and average temperature increase total biomass and
its temporal stability.

@@

```{r}
myload(qt_station, dir = mypath("data"))
matmet_cap <- paste0( 
  "Yearly temporal series of networks of two stations having respectively a low
  (A) and high (B) temporal stability of biomass. A and B stations have
  respectively biomass stability the closest to the first and third quartile of
  the distribution of the temporal stability of biomass in the dataset,
  respectively ",
  round(qt_station$bm_std_stab[1],2), " and ", round(qt_station$bm_std_stab[2],
    2),
  ". Point size are proportional to the biomass of the trophic species. (C)
  Observation sites (black dots) were located across France and monitored on the
  period 1995-2018. Red dots being the location of (A) and (B) sites. ",
  "Inner borders draw the limits of the hydrographic basins. ", "(D) The metaweb
  of the study. Each trophic species (N = 412) is represented by a node and each
  color represent a species (S = 45). Resource nodes (7) were drawn in grey. ",
  "The species names corresponding to the three digit codes are found in Table
  SXX.")

@@

# Select fishing operation 
myload(op_analysis, op_analysis_wo_holes, dir = mypath("data"))
myload(biomass_ts_sax, dir = mypath("data"))
op_analysis_bbb <- filter(op_analysis, station %in%
  biomass_ts_sax[biomass_ts_sax$sax == "bbb",]$station)
op_analysis_wo_holes_bbb <- filter(op_analysis_wo_holes, station %in%
  biomass_ts_sax[biomass_ts_sax$sax == "bbb",]$station)
range_surface <- map_dbl(c(min, max), ~.x(op_analysis_bbb$surface, na.rm =TRUE))
range_surface[1] <- ceiling(range_surface[1] / 10) * 10
range_surface[2] <- ceiling(range_surface[2] / 100) * 100
# quantile(op_analysis_bbb$surface, na.rm =TRUE)
```

@@

Structural Equation Modeling enables us to assess multivariate relationships
between variables. It is a promising way to go beyond bivariate relationships
and to investigate complex relationships, direct and
indirect, between variables [@grace_structural_2008; @grace_causal_2014;
@lefcheck_piecewisesem_2016; @grace_integrative_2016]. The structural equation
model explaining the annual total biomass allowed the environment
to affect directly species richness and network structure, as well as annual total
biomass (Figure \@ref(fig:sem-plot) B). We assumed that both species richness
and network structure directly affect total biomass (Figure \@ref(fig:sem-plot)
B) and that species richness affected foodweb structure
[@winemiller_must_1989; @dunne_network_2006; @thebault_stability_2010].

@@

library(DiagrammeR)
meta_stab_sem_rich <- export_diagram(
  fit = stab_sem_rich,
  file_path = mypath("manuscript", "bef_stability","figs", "meta_stab_sem"),
  format = "pdf",
  width = 700*2,
  height = 700*2 * asp,
  arrows = "ortho",
  val_thd = 0.05,
  env_x_pos_range = def_env_x_pos,
  env_y_pos = 0,
  env_y_sep = 1.5,
  order_env_node = c("log_RC1", "log_RC2"),
  metamodel = TRUE,
  inverse_y_node_pose = TRUE
)
p_stab_sem_rich <- export_diagram(
  fit = stab_sem_rich,
  file_path = mypath("manuscript", "bef_stability", "figs", "stab_sem"),
  format = "pdf",
  width = 700*2,
  height = 700*2 * asp,
  arrows = "ortho",
  val_thd = 0.05,
  env_x_pos_range = def_env_x_pos,
  env_y_pos = 0,
  env_y_sep = 1.5,
  order_env_node = c("log_RC1", "log_RC2"),
  inverse_y_node_pose = TRUE
)

meta_bm_sem_rich <- export_diagram(
  fit = bm_sem_rich,
  file_path = mypath("manuscript", "bef_stability", "figs", "meta_bm_sem"),
  format = "pdf",
  width = 700*2,
  height = 700*2 * asp,
  arrows = "ortho",
  val_thd = 0.05,
  env_x_pos_range = def_env_x_pos,
  env_y_pos = 0,
  env_y_sep = 1.5,
  order_env_node = c("log_RC1", "log_RC2"),
  metamodel = TRUE,
  inverse_y_node_pose = TRUE
)
p_bm_sem_rich <- export_diagram(
  fit = bm_sem_rich,
  file_path = mypath("manuscript", "bef_stability", "figs", "bm_sem"),
  format = "pdf",
  width = 700*2,
  height = 700*2 * asp,
  arrows = "ortho",
  val_thd = 0.05,
  env_x_pos_range = def_env_x_pos,
  env_y_pos = 0,
  env_y_sep = 1.5,
  order_env_node = c("log_RC1", "log_RC2"),
  inverse_y_node_pose = TRUE
)
```

@@

```{r}
sem_caption <- paste0(
  "Metamodel of the Structural Equation Models for (A) biomass stability and
  (B) total biomass. ",
  "Results of the Structural Equation Models for (C) the
  determinants of biomass stability and (D) total biomass. Red and green arrows
  represent respectively negative and positive path coefficients. The path
  coefficients are standardized. For convenience, only significant
  paths (i.e. $P < 0.05$) are shown. Rsq: Ajusted $R^2$; Avg:
  Average, CVsp: weighted temporal CV of the population; PCA1 and PCA2: respectively the first and
  second axis of the PCA."
)
```

```{r sem-plot, fig.height = 6, fig.cap = sem_caption}
#https://tex.stackexchange.com/questions/20496/drawing-different-tikz-shapes-parameterized-by-data-from-a-file?rq=1
#https://tex.stackexchange.com/questions/83888/how-to-plot-data-from-a-csv-file-using-tikz-and-csvsimple

@@

```{r indir-stab}
caption_sem <- paste0(
  "Direct, indirect and total effects of environment, species richness,
  connectance and mean trophic level on total biomass
  and its temporal stability. Indirect effects are computed by multiplying the
  path coefficients from the variable predictor to the response variable, here
  biomass stability. Indirect effects are considered as one step from the target
  variable. For exemple, the indirect effect of species richness on
  biomass stability via its effect on CVsp is $-0.62$. Only the
  significant direct effects were retained for the computation of indirect
  effects. ", "As an example, the indirect effect via species richness takes
  into account the direct effect of species richness as well as the effect of
  species richness through connectance and average trophic level.",
 "NA values indicate that the corresponding paths do not exist in the
 specified Structural Equation Model (SEM)."
)

@@

library(kableExtra)
mytable <- kable(full_table,
  format = "latex",
  booktabs = T,
  label = paste0("indir-sem"),
  caption = caption_sem
  ) %>%
#kable_paper() %>%
kable_styling(latex_options =c("scale_down")) %>%
add_header_above(c(" " = 3, "Indirect effects via" = 5)) %>%
column_spec(1:2, width = "3cm") %>%
collapse_rows(columns = 1, valign = "top")
#str_replace(mytable, "\\\\#(.*)", "\\\\label{tab:indir-sem}")
#str_view(mytable, "\\\\#(.*)")
write(mytable, file = mypath("manuscript", "bef_stability", "table.tex"))
```

@@

Surprisingly, we found a negative relationship between temporal stability and
species richness. Overall, it contrasts with experimental studies in grasslands
[@tilman_biodiversity_2006; @valencia_synchrony_2020]; mesocosms (Pennekamp et al. 2018), in 
natural communities of butterflies [@olivier_urbanization_2020] or even in lotic
fish communities [@franssen_annual_2011] where positive relationships were found
between temporal stability and species richness. However, other studies found an
opposite relationship in natural grassland [@yang_effects_2011] and a marginal
one in natural forests [@declerck_species_2006].

The first step to identify the mechanisms behind those contrasting patterns is to
look at the effects of species richness on synchrony and $\overline{CV_{sp}}$.
As most studies on experimental grasslands
[@isbell_biodiversity_2009; @roscher_identifying_2011] and natural communities
of birds, grasslands, bats and butterflies [@bluthgen_land_2016;
@olivier_urbanization_2020], we found that species richness decreases synchrony,
thereby increasing stability [but see @declerck_species_2006;
@valencia_synchrony_2020]. Our study thus extends the evidence that species
richness decreases synchrony to the case of trophic communities in natural
settings. Species richness can reduce synchrony when (i) a given increase in species
richness leads to a decrease of population sizes which raises the role of
demographic stochasticity, when (ii) species have different
responses to environmental variation [@loreau_species_2008].

The positive effect of species richness on
$\overline{CV_{sp}}$, thereby decreasing stability, is more intriguing. While
positive effect of species richness on $\overline{CV_{sp}}$ was reported in
experimental grasslands and natural bird communities [@tilman_biodiversity_2006;
@olivier_urbanization_2020], negative and neutral relationships have been found
in butterfly and bat communities. Our results contrast with a previous
meta-analysis which concludes that species richness had a positive effect on
population stability in multi-trophic aquatic microcosms
[@jiang_different_2009]. In foodwebs, the relationship between temporal
stability of populations and species richness depends on interaction strength
and its distribution [@mccann_reevaluating_1997; @thebault_trophic_2005;
@brose_allometric_2006], high interaction strength inducing a negative
relationship. Therefore, it is possible that our results are due to a high
degree of omnivory in foodwebs, which combined with strong trophic interactions
lead to more fluctuations in population biomass.

@@

In contrast, connectance had no effect on total biomass. One might have expected
that an increase in connectance may translate in a higher total nutrient uptake
and so results in higher total biomass in the fish compartment. A reason for
such an absence of an effect might be explained by the fact that our data lack
information about the lower compartments of the community, i.e., the primary
producers, zooplankton and zoobenthos.

Mean trophic level had a positive effect on biomass stability by its negative
effect on $\overline{CV_{sp}}$, thereby increasing average species stability. It
might result from a higher stability of higher trophic level as predicted by
food chain models in the presence of strong top-down control
[@barbier_pyramids_2019; @shanafelt_stability_2018]. But it
might not hold when considering omnivorous species
[@wang_biodiversity_2016]. An alternative explanation is that higher mean
trophic level leads to a stronger top-down control and resource complementarity
and thus reduce biomass fluctuations in lower trophic levels
[@rooney_structural_2006].

@@

In contrast with previous studies, we found that the average population
variability ($\overline{CV_{sp}}$) had a stronger negative effect on biomass
stability than synchrony [@tilman_biodiversity_2006;
@olivier_urbanization_2020]. It explains why we observe a negative effect of
species richness on temporal stability of biomass. The lower effect of synchrony
might be explained by the presence of generalist predators in stream
communities, that raise synchrony between prey species that share the
same predator and thus reduces compensatory
population dynamics [@raimondo_interspecific_2004]. Alternatively, it is
possible that, in aquatic foodwebs, the occurrence of large oscillatory dynamics
 leads to a higher contribution of population
variability rather than synchrony to community stability
[@mccann_reevaluating_1997]. How both strengths and the relative contribution of
synchrony and population variaribility to overall stability vary between
foodwebs and single trophic communities is still an open question.

@@

Consistent with research showing that foodwebs in wider streams support more
species and biomass [@allan_stream_2007], we found that the PCA axis related to
stream size had a positive effect on total
biomass through a positive effect on mean trophic level. Greater ecosystem size
might provide a greater diversity of resources through habitat heterogeneity,
and so higher species richness and total biomass by surface unit, through
resource complementarity. Greater diversity of resources can also maintain
more biomass at high trophic level [@post_ecosystem_2000; @doi_resource_2009].
Furthermore, the positive effect of the PCA axis related to temperature on total
biomass was of the magnitude of effects of network structure and species
richness, in line with the mounting body of work reporting important effects of
temperature on the structure and biomass of aquatic system [e.g.
@dossena_warming_2012; @edeline_ecological_2013; @emmrich_geographical_2014;
@hattab_forecasting_2016]. Contrasting with a previous study in experimental
plant communities [@grace_integrative_2016] and in fish communities
[@yvondurocher_warming_2011; @dossena_warming_2012], we found that environmental
drivers overall affect total biomass indirectly rather than directly.

Interestingly, the two PCA axis related to environmental conditions had indirect
effects on stability through two conflicting pathways, either stabilising, via
network structure, or destabilising, via species richness. The first pathway
consists in a positive effect of temperature and stream size on mean
trophic level, in line with the idea that wider streams and warmer conditions
are favourable to mesotrophic predators [@allan_stream_2007;
@emmrich_geographical_2014], which in turn stabilises biomass fluctuations,
possibly due to increased top-down regulation. The second pathway consists in a
positive effect of temperature and stream size on species richness, which itself
has a destabilising effect. This result is interesting in that both
pathways had not been evidenced to jointly play, to the best of our knowledge.
Furthermore, it is unclear which pathway is dominating, and under which circumstances,
opening an interesting avenue for further investigations.
